Proof equivalence in \capsabbrev{mll} is \capsabbrev{pspace}-complete

### 2015-06-24 Reviews LMCS/Response.txt

We thank the referees for their careful reading of our paper, and for the detailed and helpful feedback they have provided. Here, we will discuss each review in turn, as well as the changes we have made in response. We omit comments whose

Review 1

p.1, l.-1: "Canonical proof nets may also be defined for intuitionistic MLL with units": are you referring to (some of) the same papers you cite in the next paragraphs (i.e. Blute et al., Trimble, Lamarche, etc.)?  In any case, you should add a reference here.

*** We have tried and failed to find a decent reference for the kind of proof nets we have in mind here, although private discussions have convinced us that the issue is quite widely understood. The principle is simple: re-wiring jumps towards the unique conclusion yields a canonical form. But Lamarche's "essential nets", extended with units by Murawski and Ong (both of which we have added as a reference), attach jumps to the leaves of the tree instead.

p.10, notations: why do you use a "prime" in your notation for sequents/formulas?  I mean, why n' (in a circle) instead of just n?  Since n',i',j', etc. are common notations for variables (as n,i,j...), the notation is quite misleading at first.

*** The prime is used as a notation for 'successor', to indicate that a node labelled n' has n + 1 bottoms or ones, while n' and m' connect to (n+m)'.

p.25, Definition 37: I find this definition a bit disturbing, it seems ad hoc at first and leaves one wondering whether something is wrong with your encoding. [...]

*** We agree that the purpose of Definition 37 is not well-explained. We have opted for an orthogonal solution: we now explicitly define the translation of an instance of NCL-reconfiguration (which consists of a constraint graph with two configurations).

p.30, Theorem 46: just a minor quibble: you could make explicit the reduction from NCG-reconfiguration to the proof net equivalence problem, in order for the reader to better understand the complexity of such a reduction, which must be at least polytime computable. [...]

 ??? We have also made the complexity of this translation more explicit (in Theorem 46).

Review 2

I have the impression that the authors do not realize the importance of their own result.
From the point of view of the theory of proof nets, the problem concerns how to deal with lack of connectedness. It is formulated in terms of the multiplicative units but this is just for technical convenience. The reason why the problem is relevant is the fact that the bottom unit is a variant of the weakening rule, one of the cornerstones of linear logic. The multiplicative units can indeed be easily encoded by the exponentials (fix a type A and let 1 be encoded as an axiom on A, whose conclusions are connected by a par and then wrapped in a promotion, i.e. !(A^\perp \par A), and let bottom be the weakening on the dual formula). Therefore PSPACE­-completeness of proof equivalence for MLL implies PSPACE­hardness of proof equivalence for MELL. This is the real issue, as the exponentials are the expressive core of linear logic. Essentially, the paper proves that proof nets for MELL are intractable, unless ad-­hoc notions as the mix rules are added, or intractability is encapsulated in the translation from proofs to nets. Such a consequence is not explained in the paper, while there is no doubt that it has to be mentioned as it is the main reason of interest for the result.

Another issue concerns the paragraph of the introduction that mentions the fragments of linear logic for which canonical proof nets are known. I think that the polarized case should be added. Moreover it is said that Intuitionistic MLL with units has a canonical presentation but
no citations are given. Of course, *a citation has to be given*, but that is not the point. Given the results in the paper, Intuitionistic (ME)LL is the only hope for a tractable and expressive theory of proof nets. The intuitionistic case should then be discussed precisely, explaining the differences with the classical case. The canonical presentation of the intuitionistic case that the authors have in mind probably relies on jumps, and thus rewiring and proof equivalence make sense there too. Why is that case tractable? And does tractability rely on the fact that cut­-elimination in the linear case is trivial? Would Intuitionistic MELL be tractable as well?

I also find the third paragraph of page 2 (“Since there can be no canonical representation...”) unsatisfying. It points to alternative ideas in the literature to define the identity of proofs. They have to be discussed, at least briefly. At present the alternative approaches are just cited, without giving any hint of why they would help to treat the units.
I am also surprised by the total absence of a discussion of the case of MLL with the MIX rules.
A better treatment of the context of the problem cannot be avoided, because it is the reason why the result is interesting. Moreover, it is also the reason that justifies to not take too seriously the many technicalities of the

Review 3

- p.21: the bijection associates f(1) to j and c(3) to f

*** In fact, the association f(1)-f and c(3)-c in the paper is correct: each edge from parent to child is associated with the child.

### 2015-06-30 main/LMCS.tex
some intro changes

\shortJournalNames
\makeTheoremDefs

@@

@@

Proof nets were introduced at the birth of linear logic by Girard \cite{Girard-1987}, and
the correctness criterion was simplified by Danos and Regnier \cite{Danos-Regnier-1989}.
These proof nets did not include the logical units, and are canonical for multiplicative linear logic.
%
Canonical proof nets also exist for larger fragments of linear logic, such as the combined multiplicative-additive fragment without units \cite{Hughes-vanGlabbeek-2005}, and the additive fragment, including the additive units \cite{Heijltjes-2011}. Canonical proof nets may also be defined for \emph{intuitionistic} multiplicative linear logic with units.

Proof nets incorporating units were defined by Blute, Cockett, Seely and Trimble \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996}, though these proof nets cease to be canonical. The `joker games' of Koh and Ong \cite{Koh-Ong-1999} have a very similar structure to proof nets, and include units. Another explicit treatment of proof nets for \capsabbrev{mll} with units is given by Lamarche and Stra\ss burger \cite{Lamarche-Strassburger-2006}, and later a particularly simple and elegant version by Hughes \cite{Hughes-2012-categories,Hughes-2012-nets}.

@@

@@

Since there can be no canonical representation of multiplicative linear logic proofs in accordance with the categorical semantics, it is tempting to consider other ways of defining the identity of proofs. The literature contains several interesting ideas in this direction, such as \emph{polarised proof nets} \cite{Laurent-1999}, \emph{focused proof nets} \cite{Andreoli-Maieli-1999}, \emph{multi-focused proofs} \cite{Chaudhuri-Miller-Saurin-2008}, and Paul-Andr\'e Melli\`es's semantically appealing \emph{tensorial logic} and \emph{dialogue categories} \cite{Mellies-2012}.

@@

@@

An example proof net is given in Figure~\ref{fig:example net}. The edges from a $\pr$-node, which are subject to being switched, are drawn as dashed lines, while edges from a $\tn$-node are solid lines. An edge $\edge a{\links(a)}$ in a proof net or switching graph is called a \emph{link} or a \emph{jump}, and may also be indicated $\jump a{\links(a)}$, to match the illustrations. Note that unlike the axiom links of proof nets without units, a jump may connect to any connective or unit, and multiple jumps may point to the same position. A path between vertices $a$ and $b$ in a switching graph is indicated as $\path ab$.

@@

\section{Proof complexity}

@@

To encode simple numerical properties in \capsabbrev{mll}, our primary tools will be formulae of the form $\bot\tn\dotso\tn\bot$ and $1\pr\dotso\pr1$. In particular, two formulae $A$ and $B$ of the former kind may form a proof net with one $C$ of the latter kind if and only if the total number of $\bot$-occurrences in $A$ and $B$ exceeds the number of $1$-occurrences in $C$ by 1: each 1-occurrence must be linked to by a $\bot$, and one additional $\bot$ is needed to connect $A$ and $B$. We will capture this more generally by counting the number of $\bot$-occurrences in $A$ as $i+1$, so that if $j+1$ is the number of $\bot$-occurrences in $B$, the number of $1$-occurrences in $C$ must be $i+j+1$. Below left is an illustration for $i=2$ and $j=3$. In the following, we will introduce an abbreviated notation, illustrated below right. 
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\foreach \x/\i in {0/a,1/b,2/c,4/d,5/e,6/f,7/g} {\node[bullet] (\i) at (\x,0) {};}
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\draw (a)--(b)--(c) (d)--(e)--(f)--(g);
	\draw[rounded corners] ($(1) + (-.8,-.8)$) rectangle ($(6) + (.8,.8)$);
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
	\node[big_] (A) at (14,0) {$2'$};
	\node[big_] (B) at (16,0) {$3'$};
	\node[big1] (!) at (15,3) {$5'$};
	\draw[rounded corners] ($(!) + (-1.2,-.8)$) rectangle ($(!) + (1.2,.8)$);
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
By $A^n$ we will denote the sequent consisting of $n$ occurrences of a formula $A$. Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\!\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\!\pr\cdots\pr\! A_n$. In diagrammatic notation, a big disc labelled $n'$ will represent the formula $\bigtn(\bot^{n+1})$, and a big circle labelled $n'$ will represent the sequent $\1^{n+1}$. A formula $\bigparr(\1^{n+1})$ is represented by a disc enclosed in a box. This extends our diagrammatic notation as follows.

\[
\begin{array}{ccc}
	\1^{n+1}			&\Rightarrow& \bignodes1 \\[10pt]
	\bigparr(\1^{n+1})	&\Rightarrow& \bignodes2 \\[10pt]
	\bigtn(\bot^{n+1})  &\Rightarrow& \bignodes3 \\[10pt]
\end{array}
\qquad\qquad
\vc{
\begin{tikzpicture}[x=5mm,y=8mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$=$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (6,2) -- (7,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (7,0) {};
	%\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\node at (5,2) {$\dots$}; \node at (5,0) {$\dots$};
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!);
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
	\end{scope}
\end{tikzpicture}}
\qquad\quad
\begin{array}{c}
	\vc{\begin{tikzpicture}[x=5mm,y=5mm,octo] \node[big_] {$k'$}; \end{tikzpicture}} ~=~
	\vc{\begin{tikzpicture}[x=5mm,y=5mm,octo] \draw (0,0) node[big_] {$i'$} -- (2,0) node[big_] {$j'$}; \end{tikzpicture}}
	\\ \\ (k=i+j+1)
\end{array}
\]
A linking between formulae $\bigtn(\bot^n)$ and a sequent $\1^n$ will be represented by a wide arrow, as illustrated above center. For the moment, we will not distinguish between the $n!$ different ways in which such a linking can be made, but when the distinction becomes relevant we will make an explicit choice. Above right, to link a formula $\bigtn(\bot^{k+1})$ to several other formulae, we may display the formula as two discs according to the given arithmetic, representing the formula $\bigtn(\bot^{i+1})\tn~\bigtn(\bot^{j+1})$.

The primary intention behind the chosen notation is to make the arithmetic of connecting formulae $\bigtn(\bot^n)$ to formulae $\bigparr(\1^n)$ apparent in the illustrations. Specifically, the following is a proof net if and only if $n=i_1+\dotso+i_k$.
\[
  \vc{\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-2,0) {$i_1'$};
	\node[big_] (B) at ( 2,0) {$i_k'$};
	\node[big1] (!) at ( 0,3) {$n'$};
	\node at (0,0) {$\dots$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}}
\]

@@

\[
\begin{tikzpicture}[x=10mm,y=8mm,octo]
	\node[big1] (a) at (2,0) {$k'$};
	\node[big1] (b) at (5,0) {$k'$};
	\node[big1] (c) at (11,0) {$k'$};
	\draw[rounded corners] (1,-.5) rectangle (3,.5);
	\draw[rounded corners] (4,-.5) rectangle (6,.5);
	\draw[rounded corners] (10,-.5) rectangle (12,.5);
	\draw (3,0)--(4,0) (6,0)--(7,0) (9,0)--(10,0);
	%\draw[dash pattern=on 1pt off 2pt] (7.5,0)--(8.5,0) (7.5,2)--(8.5,2);
	\node at (8,0) {$\dots$}; \node at (8,2) {$\dots$};
	%
	\foreach \i in {1,...,6} {
		\node[big_] (\i) at (\i,2) {$i'_{\i}$};
	}
	\node[big_] (7) at (10,2) {};
	\node[big_] (8) at (11,2) {};
	\node[big_] (9) at (12,2) {$i'_{3n}$};
	%
	\begin{scope}[thepink,->,big>]
	\foreach \i/\a in {1/a,4/b,7/c} \draw (\i) to[bend right=20] (\a);
	\foreach \i/\a in {2/a,5/b,8/c} \draw (\i) -- (\a);
	\foreach \i/\a in {3/a,6/b,9/c} \draw (\i) to[bend left=20] (\a);
	\end{scope}
\end{tikzpicture}
\]

@@

\emph{Non-deterministic constraint graph reconfiguration} or \emph{\capsabbrev{ncg}-reconfiguration} is the problem of deciding whether two total configurations of a constraint graph are connected by a sequence of reconfiguration steps. This is the problem shown to be \capsabbrev{pspace}-complete by Hearne and Demaine \cite{GamesPuzzlesAndComputation}:

Below top is a weight element, below bottom a constraint element.
\[
\begin{array}{ccc}
	  \vc{\begin{tikzpicture}[x=5mm,y=5mm,octo]
	 	\node[big_] (A) at (0,0) {$i'$}; 
	 	\node[big_] (B) at (2,0) {$j'$};
		\node[big_] (C) at (4,0) {$k'$};
		\draw[rounded corners] (-1,-1) rectangle (5,1);
	  \end{tikzpicture}}
	&=&
	  \bigtn(\bot^{i+1})~~\pr~~\bigtn(\bot^{j+1})~~\pr~~\bigtn(\bot^{k+1})
\\ \\
	  \vc{\begin{tikzpicture}[x=5mm,y=5mm,octo]
		\node[big1] (1) at (0,0) {$m'$}; 
		\node[big1] (2) at (3,0) {$n'$};
		\draw[rounded corners] (-1,-1) rectangle (1,1) (2,-1) rectangle (4,1);
		\draw (1,0)--(2,0);
	  \end{tikzpicture}}
	&=&
	  \bigparr(\1^{m+1})~\tn~\bigparr(\1^{n+1})
\end{array}
\]
For all edges and vertices in the encoding of a constraint graph, the sum $i+j+k=m+n$ will be the same---this way, a priori any weight element may connect to any constraint element. The value $m$ will be unique for each vertex, and used for all constraint elements in the encoding of that vertex. The weight element above is for an edge connecting the two vertices for which $m=i$ and $m=i+j$: with those constraint elements it is possible to form a proof net, as illustrated below.
\[
	\connectingElements 1 \qquad \connectingElements 2
\]
To ensure that a weight element may not form a proof net with a constraint element of another vertex, the values of $m$ and $n$ are chosen such that $m\equiv1$ and $n\equiv2 \pmod 3$, and accordingly $i\equiv1$, $j\equiv0$, and $k\equiv2\pmod 3$. It then becomes impossible for $m$ to equal any other sum over $i$, $j$, and $k$, so that forming a proof net requires either $m=i$ or $m=i+j$.

An edge in a constraint graph will be encoded by an \emph{edge-gadget}, illustrated below left, constructed by stringing together a number of similar weight elements, plus a single $\bot$-formula which we will refer to as the \emph{indicator}. Illustrated below right is a \emph{vertex-gadget} encoding a vertex, formed by a number of constraint elements plus a single \emph{indicator target}.
\[
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[big_] (A) at (0,0) {$i'$}; 
	\node[big_] (B) at (2,0) {$j'$};
	\node[big_] (C) at (4,0) {$k'$};
	\draw[rounded corners] (-1,-1) rectangle (5,1);
	\node[big_] (D) at (0,4) {$i'$}; 
	\node[big_] (E) at (2,4) {$j'$};
	\node[big_] (F) at (4,4) {$k'$};
	\draw[rounded corners] (-1,3) rectangle (5,5);
	\draw (2,1)--(2,1.5) (2,2.5)--(2,3);
	\draw[dotted] (2,1.5)--(2,2.5);
	\draw (2,-1)--(2,-2) node[bullet] {};
	\path (0,-2.5)--(1,-2.5);
\end{tikzpicture}
\qquad\qquad
\begin{tikzpicture}[x=5mm,y=5mm,octo]
	\node[big1] (1) at (0,0) {$m'$}; 
	\node[big1] (2) at (3,0) {$n'$};
	\draw[rounded corners] (-1,-1) rectangle (1,1) (2,-1) rectangle (4,1);
	\draw (1,0)--(2,0);
	\node[big1] (3) at (0,4) {$m'$};
	\node[big1] (4) at (3,4) {$n'$};
	\draw[rounded corners] (-1,3) rectangle (1,5) (2,3) rectangle (4,5);
	\draw (1,4)--(2,4);
	\draw[dotted] (1.5,1.5)--(1.5,2.6);
	\node[circ] at (1.5,-2) {};
	\draw[rounded corners] (-1.5,-2.5) rectangle (4.5,5.5);
\end{tikzpicture}
\]
The natural way for edge-gadgets to link up with a vertex-gadget is shown in Figure~\ref{fig:connect example}: indicators connect to indicator targets, and weight elements connect to constraint elements or weight absorbers. The illustration depicts two equivalent proof nets, both encoding the same (fragment of a) constraint graph: a single vertex that is the target of a weight-1 edge (on the left, in red) and a weight-2 edge (on the right, in blue). The inflow of the vertex is 3, and its constraint is 2; then 2 weight elements are connected to the constraint elements in the vertex-gadget, while one remains free (connected to weight absorbers). It is essential to the functioning of the encoding that any re-distribution of weight elements gives an equivalent proof net, as illustrated by the example. Then in the second proof net, the single weight element of the left edge-gadget is free, which means the indicator jump can be rewired to connect to a different vertex.

The complete encoding of a constraint graph $G$ will then be a sequent $\Gamma$ consisting of:
%
\begin{enumerate}
 	\item
all vertex-gadgets, combined in a single formula via tensors, 
	\item
all edge-gadgets as individual formulae, and
	\item
a sufficient number of weight absorbers ($\1$-formulae).
\end{enumerate}
%
A configuration for $G$ will be encoded as a proof net for $\Gamma$, and conversely each proof net for $\Gamma$ may be interpreted as a (partial) configuration for $G$. The encoding will be made formal in Section~\ref{sec:correctness}; first we will establish some basic results for rewiring on proof nets.

@@

\noindent
In the graph for $\links$ and $\Gamma$, a box $\bx(v)$ may be \emph{closed} by replacing the subgraph over $\bx(v)$ by the single vertex $v$, and replacing every arc into $\bx(v)$ by one onto $v$. For each box $\bx(v)$ we define the \emph{local graph} to be that formed by the subgraph over $\bx(v)$ where each immediately smaller box $\bx(v)<\bx(v)$ is closed. The following is then a variation on the local retraction algorithm by Danos \cite{Danos-1990}.

@@

\begin{definition}
The \emph{empire} $\emp v$ of a vertex $v$ in $(\Gamma,\links)$ is the smallest subnet $(\Delta,\links')$ of which $v$ is a port.
\end{definition}

@@

Taking a different perspective, we may alternatively consider a switching graph as a directed tree, rooted in the root connective of $\Gamma$. This establishes a bijection between the edges and the non-root vertices, which associates each edge with its target vertex.

@@

The example in Figure~\ref{fig:directed switching graph} displays a proof net on the left, and on the right the switching graph choosing the edge $\edge ij$ for the par $i$, and the edge $\edge rg$ for the root par $r$. The induced bijection associates for example the edge named $r(1)$ with the vertex $g$, the edge $g(1)$ with $h$, and the edge $h(1)$ with $a$; it further associates $f(1)$ with $f$ and $c(3)$ with $c$.

@@

We will show that, similarly, equivalence for basic sequents is determined by parity. An immediate observation is that a proof net for a basic sequent with only one tensor-formula, every $\1$ is linked to by exactly one jump---which means that no rewiring is possible.

@@

>>>>

\longJournalNames
\makeTheoremDefs

@@

\newcommand\ap{'}

@@

Proof nets were introduced at the birth of linear logic by Girard \cite{Girard-1987}, and
the correctness criterion was simplified by Danos and Regnier \cite{Danos-Regnier-1989}.
These proof nets did not include the logical units, and are canonical for multiplicative linear logic.
%
Canonical proof nets also exist for larger fragments of linear logic, such as the combined multiplicative-additive fragment without units \cite{Hughes-vanGlabbeek-2005}, and the additive fragment, including the additive units \cite{Heijltjes-2011}.
% Canonical proof nets may also be defined for \emph{intuitionistic} multiplicative linear logic with units.

Proof nets incorporating units were defined by Blute, Cockett, Seely and Trimble \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996}, though these proof nets cease to be canonical. The `joker games' of Koh and Ong \cite{Koh-Ong-1999}, for MLL and its intuitionistic variant IMLL, have a very similar structure to proof nets, and include units. Another explicit treatment of proof nets for \capsabbrev{mll} with units is given by Lamarche and Stra\ss burger \cite{Lamarche-Strassburger-2006}, and later a particularly simple and elegant version by Hughes \cite{Hughes-2012-categories,Hughes-2012-nets}.

% NEW

The difficulty with the multiplicative units is that each instance of $\bot$, the unit to the par, must be attached to the main proof structure, while there isn't always a canonical point where to attach it. In the sequent calculus, $\bot$-instances are introduced by a weakening rule, which permutes with many (but not all) other rules. In proof nets, $\bot$-instances are commonly attached by edges called jumps, which may be re-attached via local graph-rewrite rules \cite{Trimble-1994,Blute-Cockett-Seely-Trimble-1996}. For intuitionistic MLL the rewrite rules may be directed towards the unique conclusion, to yield canonical forms.\footnote{The canonical representation for IMLL described here appears to be folklore; established proof nets for intuitionistic linear logic, called essential nets \cite{Lamarche-2008,Murawsky-Ong-2003}, (non-canonically) attach jumps to the leaves of the formula tree, for easier composition.}
For classical MLL with units however, there is not an obvious preferred point to attach jumps. And this problem extends to multiplicative--exponential linear logic (MELL) without units, where a formula $?A$ may be introduced via a weakening rule with similar permutability to the $\bot$-introduction rule.

% OLD

@@

% NEW

The result has a similar consequence for MELL without units, which may emulate the unit $\bot$ via carefully chosen formulae $?A$: equivalence of MELL proofs in normal form is \capsabbrev{pspace}-hard. This is in sharp contrast with many intuitionistic calculi such as the simply typed lambda-calculus, where normal forms are unique. However, it does not impact the complexity of proof equivalence for MELL in general, which is dominated by cut-elimination (MELL encodes the simply-typed lambda-calculus, which is not elementary recursive \cite{Statman-1977}).

@@

%Since there can be no canonical representation of multiplicative linear logic proofs in accordance with the semantics of *-autonomous categories,

Given our results for classical linear logic, which effectively rule out canonical forms for MELL, it becomes tempting to consider possible alternatives. One is intuitionistic linear logic, mentioned above. Another is Melli\`es's game-semantics inspired \emph{tensorial logic} \cite{Mellies-2012}, which embeds linear logic via a double-negation translation. Furthermore, the \emph{polarised} fragment of linear logic has canonical proof nets \cite{Laurent-1999}. A different direction to the above three, aimed more at proof search than semantics of programming, is \emph{focusing}, which retains classical linear logic but weakens the notion of proof equivalence; this gives canonical representations in \emph{focused proof nets} \cite{Andreoli-Maieli-1999} and \emph{multi-focused proofs} \cite{Chaudhuri-Miller-Saurin-2008}.

% OLD

@@

% NEW

This paper is an extended version of \cite{Heijltjes-Houston-2014}.

% OLD

@@

An example proof net is given in Figure~\ref{fig:example net}. The edges from a $\pr$-node, which are subject to being switched, are drawn as dashed lines, while edges from a $\tn$-node are solid lines. An edge $\edge a{\links(a)}$ in a proof net or switching graph is called a \emph{link} or a \emph{jump}, and may also be indicated $\jump a{\links(a)}$, to match the illustrations. Note that unlike the axiom links of proof nets without units, a jump may connect to any connective or unit, even another $\bot$-instance, and multiple jumps may point to the same position. A path between vertices $a$ and $b$ in a switching graph is indicated as $\path ab$.

@@

\section{Encoding numerical constraints}

In this section we will show how numerical constraints may be encoded in \capsabbrev{mll}, a necessary ingredient for our main construction. We will illustrate the ideas with an example: an encoding of 3-partition into the problem of \capsabbrev{mll} proof search.

@@

To encode simple numerical properties in \capsabbrev{mll}, our primary tools will be formulae of the form $\bot\tn\dotso\tn\bot$ and $1\pr\dotso\pr1$. In particular, two formulae $A$ and $B$ of the former kind may form a proof net with one $C$ of the latter kind if and only if the total number of $\bot$-occurrences in $A$ and $B$ exceeds the number of $1$-occurrences in $C$ by 1: each 1-occurrence must be linked to by a $\bot$, and one additional $\bot$ is needed to connect $A$ and $B$. We will capture this more generally by counting the number of $\bot$-occurrences in $A$ as $i+1$, so that if $j+1$ is the number of $\bot$-occurrences in $B$, the number of $1$-occurrences in $C$ must be $i+j+1$. Below left is an illustration for $i=2$ and $j=3$. In the following, we will introduce an abbreviated notation, illustrated below right. 
\[
\vc{
  \begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\foreach \x/\i in {0/a,1/b,2/c,4/d,5/e,6/f,7/g} {\node[bullet] (\i) at (\x,0) {};}
	\foreach \i in {1,...,6} {\node[circ] (\i) at (\i,3) {};}
	\draw (a)--(b)--(c) (d)--(e)--(f)--(g);
	\draw[rounded corners] ($(1) + (-.8,-.8)$) rectangle ($(6) + (.8,.8)$);
	\begin{scope}[thepink,->]
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
			\draw[bend left=10]  (d) to (3);
			\draw[bend left=10]  (e) to (4);
			\draw[bend left=10]  (f) to (5);
			\draw[bend left=10]  (g) to (6);
	\end{scope}
	\node[big_] (A) at (14,0) {$2\ap$};
	\node[big_] (B) at (16,0) {$3\ap$};
	\node[big1] (!) at (15,3) {$5\ap$};
	\draw[rounded corners] ($(!) + (-1.2,-.8)$) rectangle ($(!) + (1.2,.8)$);
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}
}
\]
By $A^n$ we will denote the sequent consisting of $n$ occurrences of a formula $A$. Given a sequent $\Gamma=A_1,\dotsc,A_n$ we will write $\bigtn\Gamma$ for $A_1\!\tn\cdots\tn A_n$, and $\bigparr\Gamma$ for $A_1\!\pr\cdots\pr\! A_n$. In diagrammatic notation, a big disc labelled $n\ap$ will represent the formula $\bigtn(\bot^{n+1})$, and a big circle labelled $n\ap$ will represent the sequent $\1^{n+1}$. A formula $\bigparr(\1^{n+1})$ is represented by a disc enclosed in a box. This extends our diagrammatic notation as follows.

\[
\begin{array}{ccc}
	\1^{n+1}			&\Rightarrow& \bignodes1 \\[10pt]
	\bigparr(\1^{n+1})	&\Rightarrow& \bignodes2 \\[10pt]
	\bigtn(\bot^{n+1})  &\Rightarrow& \bignodes3 \\[10pt]
\end{array}
\qquad\qquad
\vc{
\begin{tikzpicture}[x=5mm,y=8mm,octo]
	\node[big_] (A) at (-1,2) {};
	\node[big1] (!) at (-1,0) {};
	\node at (.5,1) {$=$};
	\draw (2,2) node[bullet] (a) {} -- (3,2) node[bullet] (b) {} -- (4,2) (6,2) -- (7,2) node[bullet] (c) {}; 
	\node[circ] (1) at (2,0) {}; \node[circ] (2) at (3,0) {}; \node[circ] (3) at (7,0) {};
	%\draw[dotted] (4,2)--(5,2) (4,0)--(5.1,0);
	\node at (5,2) {$\dots$}; \node at (5,0) {$\dots$};
	\begin{scope}[thepink,->]
			\draw[big>] (A) to (!);
			\draw[bend right=10] (a) to (1);
			\draw[bend right=10] (b) to (2);
			\draw[bend right=10] (c) to (3);
	\end{scope}
\end{tikzpicture}}
\qquad\quad
\begin{array}{c}
	\vc{\begin{tikzpicture}[x=5mm,y=5mm,octo] \node[big_] {$k\ap$}; \end{tikzpicture}} ~=~
	\vc{\begin{tikzpicture}[x=5mm,y=5mm,octo] \draw (0,0) node[big_] {$i\ap$} -- (2,0) node[big_] {$j\ap$}; \end{tikzpicture}}
	\\ \\ (k=i+j+1)
\end{array}
\]
A linking between formulae $\bigtn(\bot^n)$ and a sequent $\1^n$ will be represented by a wide arrow, as illustrated above center. For the moment, we will not distinguish between the $n!$ different ways in which such a linking can be made, but when the distinction becomes relevant we will make an explicit choice. Above right, to link a formula $\bigtn(\bot^{k+1})$ to several other formulae, we may display the formula as two discs according to the given arithmetic, representing the formula $\bigtn(\bot^{i+1})\tn~\bigtn(\bot^{j+1})$.

The primary intention behind the chosen notation is to make the arithmetic of connecting formulae $\bigtn(\bot^n)$ to formulae $\bigparr(\1^n)$ apparent in the illustrations. Specifically, the following is a proof net if and only if $n=i_1+\dotso+i_k$.
\[
  \vc{\begin{tikzpicture}[x=5mm,y=-5mm,octo]
	\node[big_] (A) at (-2,0) {$i_1\ap$};
	\node[big_] (B) at ( 2,0) {$i_k\ap$};
	\node[big1] (!) at ( 0,3) {$n\ap$};
	\node at (0,0) {$\dots$};
	\begin{scope}[thepink,->,big>]
		\draw[bend right=15] (A) to (!);
		\draw[bend left=15]  (B) to (!);
	\end{scope}
  \end{tikzpicture}}
\]

@@

\[
\begin{tikzpicture}[x=10mm,y=8mm,octo]
	\node[big1] (a) at (2,0) {$k\ap$};
	\node[big1] (b) at (5,0) {$k\ap$};
	\node[big1] (c) at (11,0) {$k\ap$};
	\draw[rounded corners] (1,-.5) rectangle (3,.5);
	\draw[rounded corners] (4,-.5) rectangle (6,.5);
	\draw[rounded corners] (10,-.5) rectangle (12,.5);
	\draw (3,0)--(4,0) (6,0)--(7,0) (9,0)--(10,0);
	%\draw[dash pattern=on 1pt off 2pt] (7.5,0)--(8.5,0) (7.5,2)--(8.5,2);
	\node at (8,0) {$\dots$}; \node at (8,2) {$\dots$};
	%
	\foreach \i in {1,...,6} {
		\node[big_] (\i) at (\i,2) {$i\ap_{\i}$};
	}
	\node[big_] (7) at (10,2) {};
	\node[big_] (8) at (11,2) {};
	\node[big_] (9) at (12,2) {$i\ap_{3n}$};
	%
	\begin{scope}[thepink,->,big>]
	\foreach \i/\a in {1/a,4/b,7/c} \draw (\i) to[bend right=20] (\a);
	\foreach \i/\a in {2/a,5/b,8/c} \draw (\i) -- (\a);
	\foreach \i/\a in {3/a,6/b,9/c} \draw (\i) to[bend left=20] (\a);
	\end{scope}
\end{tikzpicture}
\]

@@

@@

\noindent
In the graph for $\links$ and $\Gamma$, a box $\bx(v)$ may be \emph{closed} by replacing the subgraph over $\bx(v)$ by the single vertex $v$, and replacing every arc into $\bx(v)$ by one onto $v$. For each box $\bx(v)$ we define the \emph{local graph} to be that formed by the subgraph over $\bx(v)$ where each immediately smaller box $\bx(w)<\bx(v)$ is closed. The following is then a variation on the local retraction algorithm by Danos \cite{Danos-1990}.

@@

\begin{definition}
The \emph{empire} $\emp v$ of a vertex $v$ in $(\Gamma,\links)$ is the largest subnet $(\Delta,\links')$ of which $v$ is a port.
\end{definition}

@@

Taking a different perspective, we may alternatively consider a switching graph as a directed tree, rooted in the root connective of $\Gamma$. This establishes a bijection between the non-root vertices and the edges, which associates vertex with the edge connecting it to its parent.

@@

The example in Figure~\ref{fig:directed switching graph} displays a proof net on the left, and on the right the switching graph choosing the edge $\edge ij$ for the par $i$, and the edge $\edge rg$ for the root par $r$. The induced bijection associates for example the vertex $g$ and the edge named $r(1)$, the the vertex $h$ and the edge $g(1)$, and the vertex $a$ and the edge $h(1)$; it further associates $f$ with $f(1)$, and $c$ with $c(3)$.

@@

We will show that, similarly, equivalence for basic sequents is determined by parity. An immediate observation is that a proof net for a basic sequent with only one tensor-formula (a formula of the form $\bigtn(\bot^n)$ for $n\geq 2$), every $\1$ is linked to by exactly one jump, which means that no rewiring is possible.

@@

\emph{Non-deterministic constraint graph reconfiguration} or \emph{\capsabbrev{ncg}-reconfiguration} is the problem of deciding whether two total configurations of a constraint graph are connected by a sequence of reconfiguration steps. More formally, an instance of \capsabbrev{ncg}-reconfiguration is a triple $(G,\gamma,\delta)$ consisting of a constraint graph $G$ and two configurations $\gamma$ and $\delta$ for $G$. The decision problem then asks whether $\gamma\perm\delta$.

This is the problem shown to be \capsabbrev{pspace}-complete by Hearne and Demaine \cite{GamesPuzzlesAndComputation}:
